Title here

### 2020-09-01 analysis/paper/paper.tex

%% Start with one of the following:
% DOUBLE-SPACED VERSION FOR SUBMISSION TO THE AMS
\documentclass[]{ametsocV5}

@@

\usepackage{gensymb}

@@

We used monthly precipitation data from CPC Merged Analysis of
Precipitation \citep{xie1997} 2.5 degree latitude x 2.5 degree
longitude.

@@

Figure \ref{fig:asymsam-timeseries} shows the resulting Asymmetric and
Symmetric time series corresponding to 700 and 30hPa. \#FIXME

@@

The stratospheric Symmetric SAM varies strongly with a two-year period,
which can be seen using spectral methods (Figure A3) or in the
autocorrelation structure (Figure A4). There is a local peak at 2 years
in the periodigram of the tropospheric Symmetric SAM also, although it's
not statistically significant. In the troposphere the most significant
peak of variability is found in the Asymmetric index at around 3.6
months.

Correlations between the Asymmetric and Symmetric series are constant
throught the troposphere, fluctiating between 0.39 and 0.45 (Figure
\ref{fig:cor-lev}). Futhermore, the cross-correlation of each series
across levels --shown in Figure \ref{fig:cross-correlation}-- are high
in the trosposphere (greater than 0.9) for both indices. This suggests
that both the Asymmetric and the Symemtric component of the tropospheric
SAM are highly vertically coherent, both in their individual evolution
and their temporal relationship. This is to be expected since the SAM is
mostly equivalent barotropic (citaaaa).

In the stratosphere the situation is different. As can be seen in Figure
\ref{fig:cor-lev}, the relationship between the Asymmetric and Symmetric
indices varies with height above 100 hPa. It starts to decreese right
over the tropopause, reaches a minium of 0.21 at 20 hPa and then it
increases again monotonically with height up to the uppermost level of
the reanalysis. The cross-correlation across levels in the stratosphere
is generally weaker than in the troposphere (Figure
\ref{fig:cross-correlation}). Furthermore, above 100 hPa, the
cross-correlation decreases more rapdily with height for the Symmetric
SAM than for the Asymmetric SAM as evidenced by the wider dark red areas
near the diagonal in Figure \ref{fig:cross-correlation}b) vs.~Figure
\ref{fig:cross-correlation}c). Moreover, the stratospheric Symmetric SAM
seems to be slightly more connected to the trosposphere than the
Asymmetric SAM; this can be seen by the lower correlation values in the
top right quadrant of Figure \ref{fig:cross-correlation}b) in comparison
with Figure \ref{fig:cross-correlation}c).

\begin{figure}
\includegraphics{cor-lev-1} \caption[Correlation between the Symmetric and Asymmetric SAM at each level]{Correlation between the Symmetric and Asymmetric SAM at each level.}\label{fig:cor-lev}
\end{figure}

@@

Figure \ref{fig:cross-correlation}a) show the cross-correlation across
levels for the Full, Symmetric and Asymmetric SAM indices. The high
values below 100 hPa in panel reflects the vertical (zero-lag) coherency
of the tropospheric SAM. Above 100 hPa correlation between levels falls
of rapidly, indicating less coherent (zero-lag) variability. At the same
time, there is a non negligible correlation between the troposphere and
the lower-to-middle stratosphere. By examining panels b and c we see
that the Asymemtric and Symmetric SAM share the smae high level of
coherency in the tropospherem but that the Asymmetric SAM is relatively
more coherent in the stratosphere. The tropospheric-stratospheric
connection is only present in the Symmetric SAM.

\begin{figure}
\includegraphics{trends-1} \caption[Trends for each index at each level]{Trends for each index at each level. Shading indicates the 95\% confidence interval.}\label{fig:trends}
\end{figure}

Figure \ref{fig:trends} shows normalised decadal trends for each index
for the whole period along with a 95\% interval in shading. There is a
statistically significant increase in positive SAM in the tropostphere
(panel a), which has been already documented in other studies (e.g.
\citet{fogt2020}). Panels b and c of Figure \ref{fig:trends} show that
this increase is evident only in the Symmetric component. This
distinction should prove useful when attributing trends in other
variables such as temperature and precipitation to trends in the SAM.

@@

To undertand the spatial patterns associated with both indeces, we
regressed monthly geopotential anomalies into both indeces using
multiple regression (Figure A6 illustrates the difference between
computing two simple regressions and one multiple regression).

@@

Figure \ref{fig:2d-regr} shows the spatial year-long regression for
selected levels. In the troposphere the Full annular mode is clearly
``contaminated'' with well known zonal asymemtries (panels g and j)
which are succesfully sepparated by our methodology (panels h, i, k and
l). In the stratosphere, the spatial pattern associated with the Full
SAM is much more clearly dominated by a zonally symmetric, monopolar
structure (panels a and d) that is, however, not perfectly centered in
the south pole. The monopoloe obtained by multiple regression with the
Asymmetric and Symmetric SAM (panels c and f in Figure
\ref{fig:2d-regr}) is much more symmetric and the shift from total
symmetry is captured by the regression pattern of the Asymmetric SAM as
a wave-1 pattern (panels b and e).

@@

The amplitude of each zonal wave number at each latitude at 50 hPa and
700 hPa is shown in Figure \ref{fig:wave-amplitude}, where wave number
zero represents the zonal mean. Comparing between rows, this Figure
quantifies the relatively clean separation between the zonally symmetric
and zonally asymmetric structures, as its evident how the mixture of
waves of the Full field (first row) is very similar to the sum of the
waves of the Asymmetric and Symmetric field (second and third row,
respectively). The second row of Figure \ref{fig:wave-amplitude} shows
that the Asymmetric SAM is overwhelmingly dominated by wave 1 in the
stratosphere (panel b), while in the stratosphere it is composed of
zonal waves 3 to 1 in decreasing level of importance.

\begin{figure}
\includegraphics{vertical-regression-1} \caption[Asymmetric coefficient of the multiple regression of mean monthly geopotential height anomalies between 65 and 40 South]{Asymmetric coefficient of the multiple regression of mean monthly geopotential height anomalies between 65 and 40 South. (this caption needs some love)}\label{fig:vertical-regression}
\end{figure}

From Figure \ref{fig:2d-regr} it appears that the vertical structure of
the Asymmetric SAM is equivalent barotropic in the troposphere but
baroclinic in the stratosphere. Anomalies are centerd in the same
locations in the troposphere (panels h and k), but show westerly
displacement in the stratosphere (panesl b and e). This is expanded in
Figure \ref{fig:vertical-regression}, which shows a vertical crossection
of the regression coefficient corresponding to the middle row of Figure
\ref{fig:2d-regr}, area-weighted averaged between 65 and 40 degrees
South. Below 100 hPa, anomalies are completely vertical, while above
they show an important westerly tilt with height.

@@

\begin{sidewaysfigure}
\includegraphics{pp-regr-global-1} \caption[Regression pattern of precipiation with Asymmetric and Symmetric SAM]{Regression pattern of precipiation with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas.}\label{fig:pp-regr-global}
\end{sidewaysfigure}

@@

CMAP Precipitation data provided by the NOAA/OAR/ESRL PSL, Boulder,
Colorado, USA, from their Web site at https://psl.noaa.gov/

@@

\begin{figure}
\includegraphics{A5-1} \appendcaption{A5}{Regression pattern of precipiation with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas.}\label{fig:A5}
\end{figure}

@@

\begin{figure}
\includegraphics{A8-1} \appendcaption{A8}{Decadal trends of SAM indices for each season. }\label{fig:A8}
\end{figure}

>>>>

%% Start with one of the following:
% DOUBLE-SPACED VERSION FOR SUBMISSION TO THE AMS
\documentclass[twocol]{ametsocV5}

@@

\usepackage{gensymb}
\usepackage{subfig}

@@

We used monthly precipitation data from CPC Merged Analysis of
Precipitation \citep{xie1997} 2.5 degree latitude x 2.5 degree
longitude. CPCC: {[}schneider2015{]} \#FIXME

@@

Figure \ref{fig:asymsam-timeseries} shows the resulting Asymmetric and
Symmetric time series corresponding to 700 and 50hPa. \#FIXME

@@

The stratospheric Symmetric SAM varies strongly witha two-year period,
which can be seen using spectral methods (Figure A3) or in the
autocorrelation structure (Figure A4). There is a local peak at 2 years
in the periodigram of the tropospheric Symmetric SAM also, although it's
not statistically significant. In the troposphere the most significant
peak of variability is found in the Asymmetric index at around 3.6
months.

From Figure \ref{fig:asymsam-timeseries} we can see that the Asymemtric
and Symmetric time series appear to be correlated. Moreover, looking at
the extremes in the stratosphere, the Symmetric serie appears to lag the
Asymmetric series (see, for example, the positive events on late 1987
marked with a circle). We show these correlations, across all the levels
of the reanalysis and for zero and -1 lag (Asymmetric index leading the
Symmetric index), in Figure \ref{fig:cor-lev}.

\begin{figure}
\includegraphics{cor-lev-1} \caption[Correlation between the Symmetric and Asymmetric SAM at each level for lag zero and lag -1 (Asymmetric leads Symmetric)]{Correlation between the Symmetric and Asymmetric SAM at each level for lag zero and lag -1 (Asymmetric leads Symmetric).}\label{fig:cor-lev}
\end{figure}

Zero-lag correlations between the Asymmetric and Symmetric series are
relatively constant throught the troposphere, fluctiating between 0.39
and 0.45. One-month-lag correlations are similarly constant but
significantly reduced, hovering around 0.17. In the stratosphere,
zero-lag correlations drop to a minimum of 0.21 at 20 hPa and then it
increases again monotonically with height up to the uppermost level of
the reanalysis. At the same time, one-month-lag correlations increase
with height.

@@

Figure \ref{fig:cross-correlation}a) shows (zero-lag) cross-correlation
across levels for the Full, Symmetric and Asymmetric SAM indices. For
the Full SAM (panel a), high values below 100 hPa reflect the vertical
(zero-lag) coherency throughout the troposfere. Above 100 hPa
correlation between levels falls off more rapidly, indicating less
coherent (zero-lag) variability. Still there is a non negligible
correlation between the troposphere and the lower-to-middle
stratosphere. Examining panels b and c, we see that the Asymemtric and
Symmetric SAM share the same high level of coherency in the troposphere
but they differ in their stratospheric behaviour. As evidenced by the
wider dark red areas near the diagonal in Figure
\ref{fig:cross-correlation}b) vs.~Figure \ref{fig:cross-correlation}c),
stratospheric coherency is stronger for the Asymmetric SAM than the
Symemtric SAM. The stratospheric Symmetric SAM seems to connect more
strongly to the trosposphere than the Asymmetric SAM; this can be seen
by the lower correlation values in the top right left of Figure
\ref{fig:cross-correlation}b) in comparison with Figure
\ref{fig:cross-correlation}c).

\begin{figure*}
\includegraphics{trends-1} \caption[Decadal normalised trends for each index at each level for annual (row a) and seasonal values (rows b-e) for the period 1979-2018]{Decadal normalised trends for each index at each level for annual (row a) and seasonal values (rows b-e) for the period 1979-2018. Shading indicates the 95\% confidence interval.}\label{fig:trends}
\end{figure*}

Figure \ref{fig:trends} shows normalised decadal trends for each index
for the whole period 1979-2918 along with the 95\% confidence interval
in shading for the whole year (row a) and separed by trimesters (rows b
through e). As documented by \#FIXME (e.g. \citet{fogt2020}), there is a
statistically significant increase towards more positive SAM (panel
a.1), which is XX only in Summer and Autumn (panels b.1 and c.1). We
observe these increases mainly in the troposphere, reaching their
maximum at at 100 hPa in Summer. By separating the SAM signal in its
Asymmetric and Symmetric parts, we can not only see that these trends
are almost entirely due to the Symmetric component (columns 2
vs.~columns 3), but in some cases the trends become more clear. In
Summer, the Asymmetric SAM has a statistically non significant negative
trend in the middle troposphere that obscures the signal; as a result,
trends computed using only the Symmetric component are more clear
(compare the shading region in panel b.1 and b.3). In Autumn, using the
Symmetric SAM reveals a statistically significant positive trend in the
stratosphere that is not significant using the Full index.

We stress that these are only linear trends during the whole period and
the absence of a statistically significant signal should not be taken as
evidence of no sistematic change. In particular, going back to Figure
\ref{fig:asymsam-timeseries}, we can see an evident change in the
stratospheric Asymemtric component (red line in panel a) between the
90's, when we see a dominance of extreme negative values, and the 00's,
when we see the inverse. This change is restricted to the Winter months:
the linear trend for JJA starting in 1990 for the Asymmetric component
at 50hPa is \(0.37 \pm 0.22\).

These results point to very different behaviours between the
stratospheric and tropospheric SAM. \#FIXME

This suggests that both the Asymmetric and the Symemtric component of
the tropospheric SAM are highly vertically coherent, both in their
individual evolution and their temporal relationship. This is to be
expected since the SAM is mostly equivalent barotropic (citaaaa).

@@

To undertand the spatial patterns associated with both indeces, we
regressed monthly geopotential anomalies into both indeces using
multiple regression

@@

Figure \ref{fig:2d-regr} shows the spatial year-long regression \#FIXME.
Column a are regressinos using the Full SAM, while columns b and c are
regression coefficients computed in a multiple regression of
geopotential height on the Asymmetric and Symmetric indices at the same
time. Thus, they are to be interpreted as the patterns associated with
each index, controling for the (linear) effect of the other. (Figure A6
\#FIXME illustrates the difference between computing two simple
regressions and one multiple regression.)

In the stratosphere, the spatial pattern associated with the Full SAM is
more clearly dominated by a zonally symmetric, monopolar structure
(panel a.1) which is, however, not perfectly centered in the south pole.
The monopole obtained by multiple regression with the Asymmetric and
Symmetric SAM (panel c.1) is much more symmetric and the shift from
total symmetry is captured by the regression pattern of the Asymmetric
SAM as a wave-1 with maximum anomalies above the Belinghausen Sea on the
Western Hemisphere and and Davids Sea in the Eastern Hemisphere (panel
b.1).

In the troposphere, panel a.2 show the well known zonally symmetrical
annular mode \emph{contaminated} with zonal asymmetries in the form of a
wave-3. The regression using the Asymmetric and Symmetric SAM indices
successfully disentangle both structures. The Asymmetric component gives
rise to a cleaner zonal wave (panel b.2) and the Symemtric component is
assosiated with an trully annular mode, almost devoid of zonal
asymmetries (panel c.2). \#FIXME (agregar algo más?)

@@

The amplitude of each zonal wave number at each latitude at 50 hPa and
700 hPa is shown in Figure \ref{fig:wave-amplitude}, where wave number
zero represents the amplitude of the zonal mean. Comparing between rows,
this Figure quantifies the relatively clean separation between the
zonally symmetric and zonally asymmetric structures, as its evident how
the mixture of waves of the Full field (first row) is very similar to
the sum of the waves of the Asymmetric and Symmetric field (second and
third row, respectively). The second row of Figure
\ref{fig:wave-amplitude} shows that the Asymmetric SAM is overwhelmingly
dominated by wave 1 in the stratosphere (panel b), while in the
troposphere it is composed of zonal waves 3 to 1 in decreasing level of
importance.

\begin{figure}
\includegraphics{vertical-regression-1} \caption[Asymmetric coefficient of the multiple regression of mean monthly geopotential height anomalies between 65 and 40 South]{Asymmetric coefficient of the multiple regression of mean monthly geopotential height anomalies between 65 and 40 South. (\#FIXME this caption needs some love)}\label{fig:vertical-regression}
\end{figure}

To analyse the vertical structure of the geopotential anomalies
asociated with the asymetric SAM index, we show a vertical cross section
of regressions of mean geopotential height between 65ªS and 40ªS for the
50 hPa Asymmetric SAM index (panel a) and for the 700 hPa Asymmetric SAM
index (panel b) (Figure \ref{fig:vertical-regression}).

The geopotential anomalies associated with the stratospheric SAM (panel
a) are clearly constrained to the stratosphere, which underscores the
disconnect between the stratospheric and tropospheric symetric SAM. The
vertical structure this signal tilts about 60ª to the West between 100
hPa and 1 hPa, suggesting baroclinic processes and polarward transport
of heat (\#FIXME is this ok?). Interestingly, the signal in the
stratosphere maximises near 10 hPa despite using the 50 hPa index for
the regression.

The tropospheric asymmetric SAM has significant signals that extend
upwards the uppermost levels of the reanalysis. In the troposphere, the
wave-3 structure is equivalent barotropic with maximum amplitude at
roughly 250 hPa. \#FIXME moaaaaar.

Interestingly, the structures shown in Figure
\ref{fig:vertical-regression} are surprisignly robust to the choice of
index level. For any stratospheric (above 100 hPa) index, the resulting
anomalies are very similar to the wave-1 structure with maximum near 10
hPa in panel a. Conversely, for any tropospheric (below 100 hPa) index,
the result is very similar to panel b. The pattern cross-correlation
between levels of each segment of the atmosphere is greater than 0.9
(Figure A\#fixme). The patterns mainly change in amplitude. The
tropospheric pattern is maximised by the 300 hPa Asymmetric SAM index
and the stratospheric pattern increased monotonically with height.

@@

\begin{figure*}
\includegraphics{pp-regr-global-africa-1} \caption[Regression pattern of precipiation with Asymmetric and Symmetric SAM]{Regression pattern of precipiation with Asymmetric and Symmetric SAM. P-values smaller than 0.05 (controlling for Flase Detection Rate) as hatched areas. Africa}\label{fig:pp-regr-global-africa}
\end{figure*}

\begin{figure*}
\includegraphics{pp-regr-global-oceania-1} \caption[Same but for oceania]{Same but for oceania}\label{fig:pp-regr-global-oceania}
\end{figure*}

\begin{figure*}
\includegraphics{pp-regr-global-america-1} \caption[Same but for america]{Same but for america}\label{fig:pp-regr-global-america}
\end{figure*}

\subsection{Conclusiones}

This suggests that both the Asymmetric and the Symemtric component of
the tropospheric SAM are highly vertically coherent, both in their
individual evolution and their temporal relationship. This is to be
expected since the SAM is mostly equivalent barotropic (citaaaa).

@@

CMAP Precipitation data provided by the NOAA/OAR/ESRL PSL, Boulder,
Colorado, USA, from their Web site at https://psl.noaa.gov/ \#FIXME

@@

@@

\begin{figure}
\includegraphics{A8-1} \appendcaption{A8}{Pattern cross-correlation \#FIXME!}\label{fig:A8}
\end{figure}
